SMT-Boosted Security Types for Low-Level MPC

### 2025-04-26 esop25/slides.tex
checkpoint

\usepackage{mathpartir}
\usepackage{comment}

@@

@@

  \begin{block}{Reconstructive Sharing: Protocol $\prog_{\mathit{sum3}}$}
    $$
    \begin{array}{lll}
      \elab{\mesg{s1}}{2} &:=& \elab{(\secret{1} \fminus \locflip \fminus \flip{x})}{1} \\ 
      \elab{\mesg{s1}}{3} &:=& \elab{\flip{x}}{1} \\ 
      \elab{\mesg{s2}}{1} &:=& \elab{(\secret{2} \fminus \locflip \fminus \flip{x})}{2} \\ 
      \elab{\mesg{s2}}{3} &:=& \elab{\flip{x}}{2} \\ 
      \elab{\mesg{s3}}{1} &:=& \elab{(\secret{3} \fminus \locflip \fminus \flip{x})}{3} \\ 
      \elab{\mesg{s3}}{2} &:=& \elab{\flip{x}}{3}
    \end{array}
    $$
  \end{block}
  $\elab{\mesg{s1}}{2}, \elab{\mesg{s1}}{3}, \elab{\locflip}{1}$ are the \cemph{secret shares} of $\sx{1}{1}$,
    etc.
  \begin{itemize}
  \item \emph{algebraically} $\sx{1}{1} \eop \elab{\mesg{s1}}{2} + \elab{\mesg{s1}}{3} + \elab{\locflip}{1}$.
  \item \emph{probabilistically} any 2 out of three shares are in a uniform random distribution.
  \end{itemize}

@@

\begin{frame}{Using Satisfiability Modulo Finite Fields (SMFF)}

  SMFF supports equational constraint definitions $\eqs$ over a finite field $\mathcal{F}_{p}$
  with a mature solver, in the style of SMT.

  \medskip
  Overture protocols $\prog$ have a straightforward interpretation as SMFF constraints $\toeq{\prog}$ , e.g.,
  the first two lines of $\prog_{\mathit{sum3}}$ translate as:
  $$
  %\toeq{
  %  \begin{array}{c}
  %  \begin{array}{lll}
  %    \elab{\mesg{s1}}{2} &:=& \elab{(\secret{1} \fminus \locflip \fminus \flip{x})}{1} \\ 
  %    \elab{\mesg{s1}}{3} &:=& \elab{\flip{x}}{1} 
  %  \end{array}\\
  %  = \\
    \begin{array}{lll}
      \elab{\mesg{s1}}{2} &\eop& \elab{\secret{1}}{1} \fminus \elab{\locflip}{1} \fminus \elab{\flip{x}}{1}\quad \wedge \\ 
      \elab{\mesg{s1}}{3} &\eop& \elab{\flip{x}}{1} 
    \end{array}
    %\end{array}
    %}
  $$

@@

\begin{frame}{Verifying Correctness}

Letting $\prog$ be the above protocol:
$$
\begin{array}{c}
  \toeq{\prog} \wedge \ttt{out} \eop \mx{ss1}{1} + \mx{ss1}{2} + \mx{ss1}{3}\\ \models\\
  \ttt{out} \eop \sx{1}{1} \fplus \sx{2}{2} \fplus \sx{3}{3}
\end{array}
$$

@@

>>>>

\usepackage{mathpartir}
\usepackage{comment}
\usepackage{fancyvrb}
\usepackage{moreverb}

@@

\newcommand{\sumthree}{\prog_{\mathit{sum3}}}

@@

  \begin{block}{Reconstructive Sharing: Protocol $\prog_{\mathit{sum3}}$}
    $$
    \begin{array}{lll}
      \elab{\mesg{s1}}{2} &:=& \elab{(\secret{1} \fminus \locflip \fminus \flip{x})}{1} \\ 
      \elab{\mesg{s1}}{3} &:=& \elab{\flip{x}}{1} \\ 
      \elab{\mesg{s2}}{1} &:=& \elab{(\secret{2} \fminus \locflip \fminus \flip{x})}{2} \\ 
      \elab{\mesg{s2}}{3} &:=& \elab{\flip{x}}{2} \\ 
      \elab{\mesg{s3}}{1} &:=& \elab{(\secret{3} \fminus \locflip \fminus \flip{x})}{3} \\ 
      \elab{\mesg{s3}}{2} &:=& \elab{\flip{x}}{3}
    \end{array}
    $$
  \end{block}
  $\elab{\mesg{s1}}{2}, \elab{\mesg{s1}}{3}, \elab{\locflip}{1}$ are the \cemph{secret shares} of $\sx{1}{1}$,
    etc.
  \begin{itemize}
  \item \emph{algebraically} $\elab{\locflip}{1} + \elab{\mesg{s1}}{2} + \elab{\mesg{s1}}{3} \eop \sx{1}{1}$.
  \item \emph{probabilistically} any 2 out of three shares are in a uniform random distribution.
  \end{itemize}

@@

\begin{frame}{cvc5 Finite Fields Theory}

  The cvc5 Finite Fields Theory supports automatate satisfiability checking of equational
  constraint definitions $\eqs$ over prime fields $\mathcal{F}_{p}$.

  Overture protocols $\prog$ have a straightforward interpretation as SMFF constraints $\toeq{\prog}$ , e.g.,
  the first two lines of $\prog_{\mathit{sum3}}$ translate as:
  $$
  %\toeq{
  %  \begin{array}{c}
  %  \begin{array}{lll}
  %    \elab{\mesg{s1}}{2} &:=& \elab{(\secret{1} \fminus \locflip \fminus \flip{x})}{1} \\ 
  %    \elab{\mesg{s1}}{3} &:=& \elab{\flip{x}}{1} 
  %  \end{array}\\
  %  = \\
    \begin{array}{lll}
      \elab{\mesg{s1}}{2} &\eop& \elab{\secret{1}}{1} \fminus \elab{\locflip}{1} \fminus \elab{\flip{x}}{1}\quad \wedge \\ 
      \elab{\mesg{s1}}{3} &\eop& \elab{\flip{x}}{1} 
    \end{array}
    %\end{array}
    %}
 $$

 Verifiable \emph{entailment} $\eqs_1 \models \eqs_2$ means that any model of $\eqs_1$ is a model
 of $\eqs_2$. Models of $\prog$ correspond exactly to valid runs of $\prog$...

@@

\begin{frame}{Verifying Protocol Correctness}

  Entailment is a mechanism for verifying protocol properties. If an interpretation
  of a protocol entails a property, it holds in all protocol runs. 
$$
\begin{array}{c}
  \toeq{\sumthree} \wedge \ttt{out} \eop \mx{ss1}{1} + \mx{ss1}{2} + \mx{ss1}{3}\\ \models\\
  \ttt{out} \eop \sx{1}{1} \fplus \sx{2}{2} \fplus \sx{3}{3}
\end{array}
$$

Correctness of subroutines (e.g., addition and multiplication gates in encrypted circuits)
or the full protocol wrt $\idealf$ can be checked automatically.
\end{frame}

@@

\begin{frame}[fragile]{Modular Typing in Prelude}
  \small{
  \begin{verbatimtab}
SHARED(x) { (m[x]@1 + m[x]@2) }

andtablegmw(x, y, z) {
  let r11 = (r[z] + (m[x] + 1) * (m[y] + 1)) in
  ... // elided
  { r1 = r11; r2 = r10; r3 = r01; r4 = r00 }
}

andgmw(z:string, x:string, y:string) {
  let tab = andtablegmw(x,y,z) in
  m[z]@2 := OT4((m[x], m[y])@2, tab.r4, tab.r3, tab.r2, tab.r1)@1;
  m[z]@1 := r[z]@1
}
postcondition: ( m[z]@1 == r[z]@1 AND
                 m[z]@2 == (SHARED(x) * SHARED(y)) + r[z]@1) ) 
  \end{verbatimtab}
  }
\end{frame}
